---
title: "Creating histograms in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## What is a histogram?
*"A **histogram** breaks the range of values of a variable into classes and displays only the count or percent of the observations that fall into each class."* (Moore & McCabe, 2006, p.14)  
<br>

Let's create a histogram for the data set below:

<center>
87, 21, 13, 44, 88, 67, 79,  1, 99, 95,  
26, 98, 37, 59, 45, 28, 51, 57, 86, 97,  
 9, 94, 16, 22, 10, 36, 85, 60, 53, 82,  
11, 84, 78,  4, 20, 71, 72, 49, 14, 100,  
83, 31, 54, 29, 81, 27, 34, 50, 19, 17  
</center>

<br>

**1. Sort the data**  
```{r}
x = c(87, 21, 13, 44, 88, 67, 79,  1, 99, 95,
      26, 98, 37, 59, 45, 28, 51, 57, 86, 97,
       9, 94, 16, 22, 10, 36, 85, 60, 53, 82,
      11, 84, 78,  4, 20, 71, 72, 49, 14, 100,
      83, 31, 54, 29, 81, 27, 34, 50, 19, 17)
sort(x)
```

<br>

**2. Create a class interval frequency table**
```{r}
intervals = cut(x, breaks = seq(0, 100, 10))
ft = table(intervals)
t = as.data.frame(ft)
knitr::kable(t, align = "c")
```

**3. Draw a bar for each class interval with the bar height indicating the counts of values in the internal**
```{r}
# Create a variable 'x' and assign it with a set of numbers
x = c(87, 21, 13, 44, 88, 67, 79,  1, 99, 95,
      26, 98, 37, 59, 45, 28, 51, 57, 86, 97,
       9, 94, 16, 22, 10, 36, 85, 60, 53, 82,
      11, 84, 78,  4, 20, 71, 72, 49, 14, 100,
      83, 31, 54, 29, 81, 27, 34, 50, 19, 17)

# Create a histogram for the values of variable 'x' 
hist(                   # Call the hist() function
  x,                    # Apply the function to variable 'x'
  breaks = 10,          # Set the number of bins to 10. TRY different numbers.
  col = "coral",        # Set bar color as "coral". TRY "violet" or "skyblue". 
  border = "white",     # Set border color as "white". TRY "lemonchiffon".
  xlab = "Numbers",     # Label the x-axis. TRY your own label.   
  main = "A histogram") # Create a plot title. Try your own title.
```

### Common Core Mathematics Standards
CCSS.MATH.CONTENT.6.SP.B.4  
Display numerical data in plots on a number line, including dot plots, histograms, and box plots.  
<br>

CCSS.MATH.CONTENT.HSS.ID.A.1  
Represent data with plots on the real number line (dot plots, histograms, and box plots).  
<br>

### References
Moore, D. S. & McCabe, G. P. (2006). Introduction to the Practice of Statistics (5th ed.). New York: W.H. Freeman and Company. 

## Basic histograms
```{r echo=TRUE}
# Create a variable 'x' and assign it with a set of numbers
x = c(87, 21, 13, 44, 88, 67, 79,  1, 99, 95,
      26, 98, 37, 59, 45, 28, 51, 57, 86, 97,
       9, 94, 16, 22, 10, 36, 85, 60, 53, 82,
      11, 84, 78,  4, 20, 71, 72, 49, 14, 100,
      83, 31, 54, 29, 81, 27, 34, 50, 19, 17)

# Create a histogram for the values of variable 'x' 
hist(x) # Call the hist() function and apply to variable 'x'
```


## Pretty histograms

### Annotated code pattern
```{r annotated-code-pattern, exercise=TRUE}
# Create a variable 'x' and assign it with a set of numbers
x = c(87, 21, 13, 44, 88, 67, 79,  1, 99, 95,
      26, 98, 37, 59, 45, 28, 51, 57, 86, 97,
       9, 94, 16, 22, 10, 36, 85, 60, 53, 82,
      11, 84, 78,  4, 20, 71, 72, 49, 14, 100,
      83, 31, 54, 29, 81, 27, 34, 50, 19, 17)

# Create a histogram for the values of variable 'x' 
hist(                   # Call the hist() function
  x,                    # Apply the function to variable 'x'
  breaks = 10,          # Set the number of bins to 10. TRY different numbers.
  col = "coral",        # Set bar color as "coral". TRY "violet" or "skyblue". 
  border = "white",     # Set border color as "white". TRY "lemonchiffon".
  xlab = "Numbers",     # Label the x-axis. TRY your own label.
  ylab = "Counts",      # Label the y-axis. TRY your own label.
  main = "A histogram") # Create a plot title. Try your own title.
```

### Code pattern and output
```{r code-pattern}
x = c(87, 21, 13, 44, 88, 67, 79,  1, 99, 95,
      26, 98, 37, 59, 45, 28, 51, 57, 86, 97,
       9, 94, 16, 22, 10, 36, 85, 60, 53, 82,
      11, 84, 78,  4, 20, 71, 72, 49, 14, 100,
      83, 31, 54, 29, 81, 27, 34, 50, 19, 17)

hist(x, breaks = 10, col = "coral", border = "white", xlab = "Numbers", ylab = "Counts", main = "A histogram")
```

### Code template

### Comments

## How to


## Example 1: Texting by the Numbers
Rachel and her mom each asked 24 of their friends the following question: “What is the average number of texts you SEND each day?” Below are the 48 responses:

0, 2, 3, 3, 5, 5, 5, 5, 5, 5.5, 6, 6, 6, 10, 12, 13, 15, 15, 16, 20, 25, 35, 36, 70, 80, 85, 110, 130, 137, 138, 138, 140, 142, 143, 145, 150, 150, 150, 150, 150, 150, 150, 155, 162, 164, 165, 175, 275

The following R code creates a histogram of this data set:
```{r echo=TRUE}
texts = c(0, 2, 3, 3, 5, 5, 5, 5, 5, 5.5,
          6, 6, 6, 10, 12, 13, 15, 15, 16, 20,
          25, 35, 36, 70, 80, 85, 110, 130, 137, 138,
          138, 140, 142, 143, 145, 150, 150, 150, 150, 150,
          150, 150, 155, 162, 164, 165, 175, 275)

hist(texts, breaks = 15, col = "forestgreen", border = "white", xlab = "Number of texts sent each day", ylab = "Number of respondents", main = "Numbers of texts sent each day among 48 respondents")
```

Source: Secondary Math One - An Integrated Approach by the Mathematics Vision Project (https://www.mathematicsvisionproject.org/)

## Example 2: Prime numbers up to 1000

Below are the prime numbers up to 1000:

  2,   3,   5,   7,  11,  13,  17,  19,  23,  29,  31,  37,  41,
 43,  47,  53,  59,  61,  67,  71,  73,  79,  83,  89,  97, 101,
103, 107, 109, 113, 127, 131, 137, 139, 149, 151, 157, 163, 167,
173, 179, 181, 191, 193, 197, 199, 211, 223, 227, 229, 233, 239,
241, 251, 257, 263, 269, 271, 277, 281, 283, 293, 307, 311, 313,
317, 331, 337, 347, 349, 353, 359, 367, 373, 379, 383, 389, 397,
401, 409, 419, 421, 431, 433, 439, 443, 449, 457, 461, 463, 467,
479, 487, 491, 499, 503, 509, 521, 523, 541, 547, 557, 563, 569,
571, 577, 587, 593, 599, 601, 607, 613, 617, 619, 631, 641, 643,
647, 653, 659, 661, 673, 677, 683, 691, 701, 709, 719, 727, 733,
739, 743, 751, 757, 761, 769, 773, 787, 797, 809, 811, 821, 823,
827, 829, 839, 853, 857, 859, 863, 877, 881, 883, 887, 907, 911,
919, 929, 937, 941, 947, 953, 967, 971, 977, 983, 991, 997   

```{r echo=TRUE}
prime = c(2,   3,   5,   7,  11,  13,  17,  19,  23,  29,  31,  37,  41,
         43,  47,  53,  59,  61,  67,  71,  73,  79,  83,  89,  97, 101,
        103, 107, 109, 113, 127, 131, 137, 139, 149, 151, 157, 163, 167,
        173, 179, 181, 191, 193, 197, 199, 211, 223, 227, 229, 233, 239,
        241, 251, 257, 263, 269, 271, 277, 281, 283, 293, 307, 311, 313,
        317, 331, 337, 347, 349, 353, 359, 367, 373, 379, 383, 389, 397,
        401, 409, 419, 421, 431, 433, 439, 443, 449, 457, 461, 463, 467,
        479, 487, 491, 499, 503, 509, 521, 523, 541, 547, 557, 563, 569,
        571, 577, 587, 593, 599, 601, 607, 613, 617, 619, 631, 641, 643,
        647, 653, 659, 661, 673, 677, 683, 691, 701, 709, 719, 727, 733,
        739, 743, 751, 757, 761, 769, 773, 787, 797, 809, 811, 821, 823,
        827, 829, 839, 853, 857, 859, 863, 877, 881, 883, 887, 907, 911,
        919, 929, 937, 941, 947, 953, 967, 971, 977, 983, 991, 997)

hist(prime, breaks = 10, col = "skyblue", border = "white", xlab = "A histogram of prime numbers up to 1000", ylab = "Frequency", main = "Prime numbers")
```

