---
title: "Creating dot plots in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## What is a dot plot?


*"**Line plot**. A method of visually displaying a distribution of data values where each data value is shown as a dot or mark above a number line. Also known as a **dot plot**"*  --- Common Core Mathematics Standards

Let's create a dot plot for the data set below:

<center>
4, 2, 2, 5, 2, 1, 3, 2, 4, 3,   
4, 3, 1, 1, 4, 3, 5, 2, 4, 4,   
2, 1, 3, 5, 2, 4, 4, 4, 2, 4,   
3, 2, 4, 4, 2, 2, 1, 3, 1, 5,   
4, 5, 5, 5, 3, 2, 3, 4, 1, 5
</center>

<br>

**1. Sort the data**  
<center>
1, 1, 1, 1, 1, 1, 1,   
2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2,   
3, 3, 3, 3, 3, 3, 3, 3, 3,   
4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4,   
5, 5, 5, 5, 5, 5, 5, 5
</center>
<br>

**2. Count each value**
```{r}
Values = c(4, 2, 2, 5, 2, 1, 3, 2, 4, 3, 4, 3, 1, 1, 4, 3, 5, 2, 4, 4, 2, 1, 3, 5, 2, 4, 4, 4, 2, 4, 3, 2, 4, 4, 2, 2, 1, 3, 1, 5, 4, 5, 5, 5, 3, 2, 3, 4, 1, 5)
t = as.data.frame(table(Values))
knitr::kable(t, align = "c")
```
<br>

**3. Draw a number line and plot based on the frequency table**
```{r echo=FALSE}
### Visualize data on a single variable

## Create a variable `x` to store a set of numeric values
x = c(4, 2, 2, 5, 2, 1, 3, 2, 4, 3, 4, 3, 1, 1, 4, 3, 5, 2, 4, 4, 2, 1, 3, 5, 2, 4, 4, 4, 2, 4, 3, 2, 4, 4, 2, 2, 1, 3, 1, 5, 4, 5, 5, 5, 3, 2, 3, 4, 1, 5)  # `c()` combines values inside the parentheses into a vector; `=` assignment operator assign the vector to variable `x`
                                   
## Create a dot plot for the values in variable `x`
stripchart(              # Call the `stripchart()` function.
  x,                     # Apply the function to variable `x`.
  method = "stack",      # Stack up points of the same value.
  frame.plot = FALSE,    # Remove the plot frame.
  at = 0,                # Position the points at the bottom.
  offset = 0.5,          # Separate stacking points by 0.5 of point size. TRY 1 or 2.
  pch = 16,              # Set point symbol as 16-solid circle. TRY 0-square or 8-star.
  col = "black",   # Set point color as deepskyblue. TRY "forestgreen" or "deeppink".
  main = "My dot plot",  # Create a plot title. TRY your own title. 
  xlab = "Values")       # Label the x-axis. TRY your own label.
```

[CCSS.MATH.CONTENT.6.SP.B.4](http://www.corestandards.org/Math/Content/6/SP/B/4/)  
Display numerical data in plots on a number line, including dot plots, histograms, and box plots.

[CCSS.MATH.CONTENT.HSS.ID.A.1](http://www.corestandards.org/Math/Content/HSS/ID/A/1/)  
Represent data with plots on the real number line (dot plots, histograms, and box plots).

## Basic dot plots
```{r basic, echo=TRUE, exercise=TRUE}
## Create a variable `x` to store a set of numeric values
x = c(4, 2, 2, 5, 2, 1, 3, 2, 4, 3, 4, 3, 1, 1, 4, 3, 5, 2, 4, 4, 2, 1, 3, 5, 2, 4, 4, 4, 2, 4, 3, 2, 4, 4, 2, 2, 1, 3, 1, 5, 4, 5, 5, 5, 3, 2, 3, 4, 1, 5)
## Create a dot plot for the values in variable `x`
stripchart(              # Call the `stripchart()` function.
  x,                     # Apply the function to variable `x`.
  method = "stack")      # Stack up points of the same value.
```

## Pretty dot plots

### Annotated code pattern
1. Press 'Run Code' on the top right to see the plot.
2. Try different argument values as suggested.
3. Press 'Start Over' on the top left to restore the default.
```{r annotated-code-pattern, echo=TRUE, exercise=TRUE}
### Visualize data on a single variable

## Create a variable `x` to store a set of numeric values
x = c(4, 2, 2, 5, 2, 1, 3, 2, 4, 3, 4, 3, 1, 1, 4, 3, 5, 2, 4, 4, 2, 1, 3, 5, 2, 4, 4, 4, 2, 4, 3, 2, 4, 4, 2, 2, 1, 3, 1, 5, 4, 5, 5, 5, 3, 2, 3, 4, 1, 5)  # `c()` combines values inside the parentheses into a vector; `=` assignment operator assign the vector to variable `x`
                                   
## Create a dot plot for the values in variable `x`
stripchart(              # Call the `stripchart()` function.
  x,                     # Apply the function to variable `x`.
  method = "stack",      # Stack up points of the same value.
  frame.plot = FALSE,    # Remove the plot frame.
  at = 0,                # Position the points at the bottom.
  offset = 0.5,          # Separate stacking points by 0.5 of point size. TRY 1 or 2.
  pch = 16,              # Set point symbol as 16-solid circle. TRY 0-square or 8-star.
  col = "deepskyblue",   # Set point color as deepskyblue. TRY "forestgreen" or "deeppink".
  main = "My dot plot",  # Create a plot title. TRY your own title. 
  xlab = "Values")       # Label the x-axis. TRY your own label.
```

### Code pattern and output
```{r code-pattern, echo=TRUE}
x = c(4, 2, 2, 5, 2, 1, 3, 2, 4, 3, 4, 3, 1, 1, 4, 3, 5, 2, 4, 4, 2, 1, 3, 5, 2, 4, 4, 4, 2, 4, 3, 2, 4, 4, 2, 2, 1, 3, 1, 5, 4, 5, 5, 5, 3, 2, 3, 4, 1, 5)
stripchart(x, method = "stack", frame.plot = FALSE, at = 0, offset = 0.5, pch = 16, col = "deepskyblue", main = "My dot plot", xlab = "Possible values")
```

### Code template
```{r code-template, echo=TRUE, eval=FALSE}
x = c(__ , __ , __)
stripchart(x, method = "stack", frame.plot = FALSE, at = 0, offset = 0.5, pch = __, col = __, main = "__", xlab = "__")
```

### Comments
```{r comments, echo=TRUE}
### Visualize data on a single variable

## Create a variable `x` to store a set of numeric values
          # `c()` combines values inside the parentheses into a vector; `=` assignment operator assign the vector to variable `x`
                                   
## Create a dot plot for the values in variable `x`
          # Call the `stripchart()` function.
          # Apply the function to variable `x`.
          # Atack up points of the same value.
          # Remove the plot frame.
          # Position the points at the bottom.
          # Separate stacking points by 0.5 of point size. TRY 1 or 2.
          # Set point symbol as 16-solid circle. TRY 0-square or 8-star.
          # Set point color as deepskyblue. TRY "forestgreen" or "deeppink".
          # Create a plot title. TRY your own title. 
          # Label the x-axis. TRY your own label.
```

## Tutorial


## Example 1: Phones per household 
### Number of mobile phones per household
<center>
2, 2, 3, 3, 2, 4, 4, 5, 4, 5, 5, 5, 6, 4, 4, 4, 4, 5, 5, 6, 4, 3, 3, 4, 3
</center>

```{r phones, echo=TRUE, exercise=TRUE}
phones = c(2, 2, 3, 3, 2, 4, 4, 5, 4, 5, 5, 5, 6, 4, 4, 4, 4, 5, 5, 6, 4, 3, 3, 4, 3)
stripchart(phones, method = "stack", frame.plot = FALSE, at = 0, offset = 1, pch = 8, col = "red", main = "Number of mobile phones per household", xlab = NULL)
```

## Example 2: Beverage prices

Prices of beverages at a coffee shop
<center>
2.25, 1.95, 2.25, 2.50, 3.50, 3.50, 4.00, 4.00, 3.25, 3.25, 4.25, 4.75, 2.75, 3.75, 4.75, 4.75, 5.00, 4.00, 5.00, 5.00
</center>
```{r beverage-prices, echo=TRUE, exercise=TRUE}
beverage = c(2.25, 1.95, 2.25, 2.50, 3.50, 3.50, 4.00, 4.00, 3.25, 3.25, 4.25, 4.75, 2.75, 3.75, 4.75, 4.75, 5.00, 4.00, 5.00, 5.00)
stripchart(beverage, method = "stack", frame.plot = FALSE, at = 0, offset = 1, pch = 4, col = "brown", main = "Prices of beverages at a coffee shop", xlab = NULL)
```

## Example 3: History test scores

A high school history teacher recorded the test scores for 30 students on a recent exam. The score are listed below:

88, 90, 85, 83, 89, 85, 81, 85, 86, 88, 85, 90, 85, 81, 86, 82, 84, 82, 83, 95, 88, 91, 92, 86, 82, 81, 81, 75, 83, 84

Using the data above, construct a dotplot to better visualize and understand the data.

```{r history-scores, echo=TRUE, exercise=TRUE}
scores = c(88, 90, 85, 83, 89, 85, 81, 85, 86, 88, 85, 90, 85, 81, 86, 82, 84, 82, 83, 95, 88, 91, 92, 86, 82, 81, 81, 75, 83, 84)
stripchart(scores, method = "stack", frame.plot = FALSE, at = 0, offset = 1, pch = 16, col = "orange", main = "History test scores", xlab = "Scores")
```

## Your scripting area


```{r yours, exercise = TRUE}














```

